---
title: "Lab notebook, setup hydrolytic enzymes"
author: "Karolina Jörgensen"
format: html
editor: visual
---

## 2023-09-05 Standard curves

Prepared commercial acid phosphatase, MUB solution, NaOH and aP substrate. I used dH<sub>2</sub>O for the enzyme and sodium acetate buffer (pH = 5) that Erica had prepared the day before for the other reagents.

Acid phosphatase, stock solution: 20 mg in 10 ml -\> 10 U/ml.\
Made a standard curve (10-fold) with 1; 0.1; 0.01; 0.001: 0.0001, 0 U/ml

MUB, stock solution: 8.8085 mg in 50 ml -\> 1 mM\
Made a standard curve with 100; 90; 60; 30; 9; 3; 1;0 µM

aP substrate 5.12 mg in 100 ml -\> 200 µM, frozen in aliquots of 5 or 10 ml (enough for 1 or 2 96 well plates)

-   Prepared a plate with 4 technical replicates of:\
    MUB standard + acetate buffer (high to low, A-H, 1-4). 50 µl MUB + 200 µl acetate buffer

-   aP enzyme + MUB standard (high to low, A-F, 5-8). 200 µl aP + 50 µl MUB

-   aP enzyme + aP substrate (high to low, A-F, 9-12). 200 µl aP + 50 µl aP substrate

Plate was incubated in darkness for 15 minutes and stopped with 10 µl NaOH.

```{r}
#| echo: false

library(tidyverse)
library(readxl)
library(ggpmisc)
# Load raw data from standard curve test
# Only read range with fluorescence data
standard1 <- read_xlsx("Data/Raw/MUB_hydrolytic_enzymes_20230905_134336.xlsx", range ="A44:M52")

# Read plate layout
standard1.samples <- read_xlsx("Data/Raw/sample_template_20230905_134336.xlsx", range = "A1:M9")
standard1.treats <- read_xlsx("Data/Raw/sample_template_20230905_134336.xlsx", range = "A12:M20")

# Wrangle sample and treatment data with "well_ID" as identifyer
standard1.samples <- standard1.samples |> 
  pivot_longer(cols = 2:13, names_to = "column", values_to = "sample_conc") |> 
  unite(well_ID, c(x, column), sep = "", remove = TRUE)

standard1.treats <- standard1.treats |> 
  pivot_longer(cols = 2:13, names_to = "column", values_to = "treatment") |> 
  unite(well_ID, c(x, column), sep = "", remove = TRUE)

standard1.samples <- full_join(standard1.samples, standard1.treats, by = "well_ID")

# Wrangle fluorescence data into long format
standard1 <- standard1 |> 
  pivot_longer(cols = 2:13, names_to = "column", values_to = "fluorescence") |> 
  unite(well_ID, c("<>", column), sep = "", remove = TRUE)

# Generate cleaned dataset  
standard1.clean <- full_join(standard1, standard1.samples, by = "well_ID") |> 
  drop_na() 

# Summarize data (averages of technical quadruplicates)
standard1.sum <- standard1.clean |> 
  group_by(treatment, sample_conc) |> 
  summarise(mean = mean(fluorescence))

# Remove highest aP concentration (range much above the MUB standard curve)

standard1.sum <- standard1.sum |> 
  filter(!(treatment == "aP_aPsub" & sample_conc == 1.0))

# Plot standard curves
standard.curve.1 <- ggplot(standard1.sum, aes(x = sample_conc, y = mean))+
  geom_point() +
  facet_grid(.~treatment, scales = "free")+
  geom_smooth(method = "lm")+
  stat_poly_eq(use_label(c("R2")))
  
```

```{r}
standard.curve.1
```

## 2023-09-14 Planning soil trials

The method worked on the new plate reader, and standard curves looked really good. Next step is to run tests with aP-spiked soil to make sure the assay works well in real soil slurries too. Should be quite easy to do!

Step 1 is preparing soil slurries. I can use the old sodium acetate that Erica prepared for the trials. The same slurry can then be used to try out the MnP protocol. Use two of the soils that have been stored in the basement - also check if Dorte has some leftover agricultural soil I could use.

Spike the soils with an appropriate amount of aP (maybe around 0.05 U/ml). Run tests on background activities in the soils + quenching control.
